---
title: "Load & Review Data v2 (get all Bone sorted)"
author: "Wesley"
date: "2022/03/28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load all of the Bone Data

To start, we need to load all of the Excel files. To do this, we will use the
readxl package. In addition, we're going to need to manipulate strings in 
the following, so let's pre-emptively load that package.

```{r libraries, error = FALSE, message = FALSE, warning = FALSE}
library("readxl")
library("stringr")
```

### Start with Bone 1, as Template

Bone 1 has many sheets, each of which is a nicely organized listing of VOCs
as detected by the GCxGC system's built-in scanning software, and exported. 
We will load them all. 

```{r load_data}
sheets <- excel_sheets("/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone1.xlsx")  # note: spaces in names
Bone1 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone1.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(Bone1) <- sheets
```

Having things organized by sample is a good thing, but we really will
find it annoying finding things long-term. In the following, I will take each
of the lists of compounds relevant to Bone 1, and merge them into a single
data.frame object, so we can then filter that frame down to get a specific
compound, or a specific day/observation, and so on.


```{r merge_sheets}
Bone1m <- do.call("rbind", Bone1)
names(Bone1m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "1stRT","2ndRT","Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

A quick print of the data structure shows that the RTs, QuantMass and BaseMass
won't be usable as numbers. The RTs need to be split, as well.

```{r split_RTs}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(Bone1m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(Bone1m$RTs, ","), FUN = "[[", 2)
         )
       )
Bone1m <- data.frame(Bone1m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples}
unique(Bone1m$Sample)
save(file = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone1.rda", Bone1m)
```

### Now Do Bone 2

```{r load_data2}
sheets <- excel_sheets("/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone2.xlsx")  # note: spaces in names
Bone2 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone2.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(Bone2) <- sheets

Bone2m <- do.call("rbind", Bone2)
names(Bone2m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs","1stRT","2ndRT", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs2}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(Bone2m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(Bone2m$RTs, ","), FUN = "[[", 2)
         )
       )
Bone2m <- data.frame(Bone2m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(Bone2m$Sample)
save(file = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone2.rda", Bone2m)
```

### Now Do Bone 3

```{r load_data3}
sheets <- excel_sheets("/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone3.xlsx")  # note: spaces in names
Bone3 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone3.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(Bone3) <- sheets

Bone3m <- do.call("rbind", Bone3)
names(Bone3m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs","1stRT","2ndRT", "Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs3}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(Bone3m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(Bone3m$RTs, ","), FUN = "[[", 2)
         )
       )
Bone3m <- data.frame(Bone3m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(Bone3m$Sample)
save(file = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone3.rda", Bone3m)
```

### Now Do Bone 4

```{r load_data4}
sheets <- excel_sheets("/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone4.xlsx")  # note: spaces in names
Bone4 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone4.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(Bone4) <- sheets

Bone4m <- do.call("rbind", Bone4)
names(Bone4m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "1stRT","2ndRT","Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(Bone4m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(Bone4m$RTs, ","), FUN = "[[", 2)
         )
       )
Bone4m <- data.frame(Bone4m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(Bone4m$Sample)
save(file = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone4.rda", Bone4m)
```

### Now Do Bone 5

```{r load_data5}
sheets <- excel_sheets("/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone5.xlsx")  # note: spaces in names
Bone5 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone5.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(Bone5) <- sheets

Bone5m <- do.call("rbind", Bone5)
names(Bone5m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "1stRT","2ndRT","Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(Bone5m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(Bone5m$RTs, ","), FUN = "[[", 2)
         )
       )
Bone5m <- data.frame(Bone5m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(Bone5m$Sample)
save(file = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone5.rda", Bone5m)
```

### Now Do Bone 6

```{r load_data6}
sheets <- excel_sheets("/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone6.xlsx")  # note: spaces in names
Bone6 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone6.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(Bone6) <- sheets

Bone6m <- do.call("rbind", Bone6)
names(Bone6m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "1stRT","2ndRT","Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(Bone6m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(Bone6m$RTs, ","), FUN = "[[", 2)
         )
       )
Bone6m <- data.frame(Bone6m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(Bone6m$Sample)
save(file = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone6.rda", Bone6m)
```

### Now Do Bone 7

```{r load_data7}
sheets <- excel_sheets("/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone7.xlsx")  # note: spaces in names
Bone7 <- lapply(sheets, FUN = function(x) {
           read_xlsx(path = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone7.xlsx",
                     sheet = x) })
sheets <- str_replace_all(sheets, " ", "_")
names(Bone7) <- sheets

Bone7m <- do.call("rbind", Bone7)
names(Bone7m) <- c("PeakNum", "Sample", "Name", "Formula", "RTs", "1stRT","2ndRT","Similarity",
                   "Area", "Height", "QuantMass", "BaseMass", "QuantSN", "PeakSN",
                   "Type", "Conc", "ConcUnits")
```

```{r split_RTs4}
RT1 <- as.numeric(
         unlist(
           lapply(strsplit(Bone7m$RTs, ","), FUN = "[[", 1)
         )
       )
RT2 <- as.numeric(
         unlist(
           lapply(strsplit(Bone7m$RTs, ","), FUN = "[[", 2)
         )
       )
Bone7m <- data.frame(Bone7m[, c(1:4, 6:17)], RT1 = RT1, RT2 = RT2)
```

```{r samples2}
unique(Bone7m$Sample)
save(file = "/home/wburr/Rushali_Analysis/New_Stuff/Bone/Bone7.rda", Bone7m)
```

