---
title: "Setting up for PCA"
author: "Wesley Burr"
date: "29/03/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
```


```{r function}
##
#  clean_common: Function to select, filter, clean, filter and merge 
#  compounds using the logic of:
#  * unique to the samples, not the control, for ALL samples; or
#  * in both samples and control, but much stronger in the control
# 
#  Inputs:
#  * dat: data.frame sourced from merging spreadsheets of GCxGC output
#  * sample_names: names of specific samples (e.g., D1ad0)
#  * control_names: names of specific control samples (e.g., Control_D1d0)
#  * ratio_par: cut-off for the logic of "in both samples and control" - if this
#      is set very large, will eliminate cross-overs.
#
#  Returns:
#  * samples_keep: list of full data.frames for individual replicates, cleaned
#      down to relevant compounds using above logic
##
clean_common <- function(dat,
                         sample_names,
                         control_names,
                         ratio_par = 2.0) {
  
  samples <- vector("list", length = length(sample_names))
  names(samples) <- sample_names
  controls <- vector("list", length = length(control_names))
  names(controls) <- control_names
  
  # Extract specific samples and controls of interest and
  # dump all but the largest Area example of each compound
  for(j in 1:length(sample_names)) {
    samples[[j]] <- dat %>% subset(Sample == sample_names[j]) %>%
                      group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() %>% filter(substr(Name, 1, 4) != "Peak")
    samples[[j]] <- samples[[j]][!duplicated(samples[[j]]$Name), ]
  } 
  for(j in 1:length(control_names)) {
    controls[[j]] <- dat %>% subset(Sample == control_names[j]) %>%
                      group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() %>% filter(substr(Name, 1, 4) != "Peak")
    controls[[j]] <- controls[[j]][!duplicated(controls[[j]]$Name), ]
  } 
  # merge controls
  control <- do.call("rbind", controls)
  control <- control %>% group_by(Name) %>%
                      filter(Area == max(Area)) %>%
                      ungroup() 
  control <- control[!duplicated(control$Name), ]

  # Find compounds that are in each sample that are also in control
  samples_keep <- samples
  for(j in 1:length(sample_names)) {
    samp <- samples[[j]] %>% filter(samples[[j]]$Name %in% control$Name)
    cont <- control %>% filter(control$Name %in% samples[[j]]$Name)
   
    # ratio is high enough to keep 
    samp_SN <- unlist(samp[order(samp$Name), "PeakSN"])
    cont_SN <- unlist(cont[order(cont$Name), "PeakSN"])
    contrib1 <- samp %>% filter((samp_SN / cont_SN) > ratio_par)
   
    # also, compounds that are *not* in the controls 
    contrib2 <- samples[[j]] %>% filter(!(samples[[j]]$Name %in% control$Name))
    samples_keep[[j]] <- rbind(contrib1, contrib2)
  }
  names(samples_keep) <- sample_names
  samples_keep
}


##
#
#  join_common: Function which takes output of clean_common above,
#    and merges based on common presence across all replicates of compounds. 
#
#  Inputs:
#  * compounds: list of data.frames, 16 columns as in the spreadsheets
#  
#  Outputs:
#  * common: merged, simplified data.frame, created via inner_join of data.frames after filtering.
##
join_common <- function(compounds) {
  n_samp <- length(compounds)
  subset_compounds <- vector("list", length = n_samp)
  for(j in 1:n_samp) {
    subset_compounds[[j]] <- compounds[[j]]
    if(n_samp > 1) {
      for(k in (1:n_samp)[-j]) {
        subset_compounds[[j]] <- subset_compounds[[j]] %>%
                                   subset(subset_compounds[[j]]$Name %in% compounds[[k]]$Name)
      }
    }
    subset_compounds[[j]] <- subset_compounds[[j]] %>% select(Name, Area, PeakSN)
  }
  
  # Join first two, if they exist
  if(n_samp > 1) {
    common <- inner_join(x = subset_compounds[[1]], y = subset_compounds[[2]], by = "Name")
    if(n_samp >= 3) {
      for(j in 3:n_samp) {
        common <- inner_join(x = common, y = subset_compounds[[j]], by = "Name")  
      }
    }
  } else {
    common <- subset_compounds[[1]][, c("Name", "Area", "PeakSN")]
  }
  names(common) <- c("Name", paste0(c("Area_", "PeakSN_"), rep(1:n_samp, each = 2)))
  common
}
```

# Working Through Donor

```{r Donor2}
load("/home/wburr/Rushali_REST/Donor2/D2n.rda")
D2m$PeakSN<-as.numeric(D2m$PeakSN)
D2m$Sample <- str_replace_all(D2m$Sample, " ", "_")

D2d0 <- join_common( clean_common(D2m,
                sample_names = c("D2ad0", "D2bd0", "D2cd0"),
                control_names = c("Control_D2d0"),
                ratio_par = 2.0) )
D2d1 <- join_common( clean_common(D2m,
                sample_names = c("D2ad1", "D2bd1", "D2cd1"),
                control_names = c("Control_D2d1"),
                ratio_par = 2.0) )
D2d2 <- join_common( clean_common(D2m,
                sample_names = c("D2ad2", "D2bd2", "D2cd2"),
                control_names = c("Control_D2d2"),
                ratio_par = 2.0) )
D2d4 <- join_common( clean_common(D2m,
                sample_names = c("D2ad4", "D2bd4", "D2cd4"),
                control_names = c("Control_D2d4"),
                ratio_par = 2.0) )
D2d5 <- join_common( clean_common(D2m,
                sample_names = c("D2ad5", "D2bd5", "D2cd5"),
                control_names = c("Control_D2d5"),
                ratio_par = 2.0) )
D2d6 <- join_common( clean_common(D2m,
                sample_names = c("D2ad6", "D2bd6", "D2cd6"),
                control_names = c("Control_D2d6"),
                ratio_par = 2.0) )
D2d7 <- join_common( clean_common(D2m,
                sample_names = c("D2ad7", "D2bd7", "D2cd7"),
                control_names = c("Control_D2d7"),
                ratio_par = 2.0) )
D2d8 <- join_common( clean_common(D2m,
                sample_names = c("D2ad8"),
                control_names = c("Control_D2d8"),
                ratio_par = 2.0) )
D2d9 <- join_common( clean_common(D2m,
                sample_names = c("D2ad9", "D2bd9"),
                control_names = c("Control_D2d9"),
                ratio_par = 2.0) )
D2d10 <- join_common( clean_common(D2m,
                sample_names = c("D2ad10", "D2bd10", "D2cd10"),
                control_names = c("Control_D2d10"),
                ratio_par = 2.0) )
D2d11 <- join_common( clean_common(D2m,
                sample_names = c("D2ad11", "D2bd11", "D2cd11"),
                control_names = c("Control_D2d11"),
                ratio_par = 2.0) )
D2d12 <- join_common( clean_common(D2m,
                sample_names = c("D2ad12", "D2bd12", "D2cd12"),
                control_names = c("Control_D2d12"),
                ratio_par = 2.0) )
D2d13 <- join_common( clean_common(D2m,
                sample_names = c("D2ad13", "D2bd13", "D2cd13"),
                control_names = c("Control_D2d13"),
                ratio_par = 2.0) )
D2d14 <- join_common( clean_common(D2m,
                sample_names = c("D2ad14", "D2bd14", "D2cd14"),
                control_names = c("Control_D2d14"),
                ratio_par = 2.0) )
D2d15 <- join_common( clean_common(D2m,
                sample_names = c("D2ad15", "D2bd15", "D2cd15"),
                control_names = c("Control_D2d15"),
                ratio_par = 2.0) )
D2d16 <- join_common( clean_common(D2m,
                sample_names = c("D2ad16", "D2bd16", "D2cd16"),
                control_names = c("Control_D2d16"),
                ratio_par = 2.0) )
D2d17 <- join_common( clean_common(D2m,
                sample_names = c("D2ad17", "D2bd17", "D2cd17"),
                control_names = c("Control_D2d17"),
                ratio_par = 2.0) )
D2d18 <- join_common( clean_common(D2m,
                sample_names = c("D2ad18", "D2bd18", "D2cd18"),
                control_names = c("Control_D2d18"),
                ratio_par = 2.0) )
D2d21 <- join_common( clean_common(D2m,
                sample_names = c("D2ad21", "D2bd21", "D2cd21"),
                control_names = c("Control_D2d21"),
                ratio_par = 2.0) )
D2d26 <- join_common( clean_common(D2m,
                sample_names = c("D2ad26", "D2bd26", "D2cd26"),
                control_names = c("Control_D2d26"),
                ratio_par = 2.0) )
D2d28 <- join_common( clean_common(D2m,
                sample_names = c("D2ad28", "D2bd28", "D2cd28"),
                control_names = c("Control_D2d28"),
                ratio_par = 2.0) )
D2d30 <- join_common( clean_common(D2m,
                sample_names = c("D2ad30", "D2bd30", "D2cd30"),
                control_names = c("Control_D2d30"),
                ratio_par = 2.0) )
D2d32 <- join_common( clean_common(D2m,
                sample_names = c("D2ad32", "D2bd32", "D2cd32"),
                control_names = c("Control_D2d32"),
                ratio_par = 2.0) )
D2d35 <- join_common( clean_common(D2m,
                sample_names = c("D2ad35", "D2bd35", "D2cd35"),
                control_names = c("Control_D2d35"),
                ratio_par = 2.0) )
D2d42 <- join_common( clean_common(D2m,
                sample_names = c("D2ad42", "D2bd42", "D2cd42"),
                control_names = c("Control_D2d42"),
                ratio_par = 2.0) )
D2d49 <- join_common( clean_common(D2m,
                sample_names = c("D2ad49", "D2bd49", "D2cd49"),
                control_names = c("Control_D2d49"),
                ratio_par = 2.0) )
D2d57 <- join_common( clean_common(D2m,
                sample_names = c("D2ad57", "D2bd57", "D2cd57"),
                control_names = c("Control_D2d57"),
                ratio_par = 2.0) )
D2d65 <- join_common( clean_common(D2m,
                sample_names = c("D2ad65", "D2bd65", "D2cd65"),
                control_names = c("Control_D2d65"),
                ratio_par = 2.0) )
D2d74 <- join_common( clean_common(D2m,
                sample_names = c("D2ad74", "D2bd74", "D2cd74"),
                control_names = c("Control_D2d74"),
                ratio_par = 2.0) )
D2d81 <- join_common( clean_common(D2m,
                sample_names = c("D2ad81", "D2bd81", "D2cd81"),
                control_names = c("Control_D2d81"),
                ratio_par = 2.0) )
D2d87 <- join_common( clean_common(D2m,
                sample_names = c("D2ad87", "D2bd87", "D2cd87"),
                control_names = c("Control_D2d87"),
                ratio_par = 2.0) )
```




```{r}
all_d <- vector("list", 31)
all_d[[1]] <- D2d0
all_d[[2]] <- D2d1
all_d[[3]] <- D2d2
all_d[[4]] <- D2d4
all_d[[5]] <- D2d5
all_d[[6]] <- D2d6
all_d[[7]] <- D2d7
all_d[[8]] <- D2d8
all_d[[9]] <- D2d9
all_d[[10]] <- D2d10
all_d[[11]] <- D2d11
all_d[[12]] <- D2d12
all_d[[13]] <- D2d13
all_d[[14]] <- D2d14
all_d[[15]] <- D2d15
all_d[[16]] <- D2d16
all_d[[17]] <- D2d17
all_d[[18]] <- D2d18
all_d[[19]] <- D2d21
all_d[[20]] <- D2d26
all_d[[21]] <- D2d28
all_d[[22]] <- D2d30
all_d[[23]] <- D2d32
all_d[[24]] <- D2d35
all_d[[25]] <- D2d42
all_d[[26]] <- D2d49
all_d[[27]] <- D2d57
all_d[[28]] <- D2d65
all_d[[29]] <- D2d74
all_d[[30]] <- D2d81
all_d[[31]] <- D2d87

names(all_d) <- c(paste0("D2d0"),
                  paste0("D2d1"),
                  paste0("D2d2"),
                  paste0("D2d4"),
                  paste0("D2d5"),
                  paste0("D2d6"),
                  paste0("D2d7"),
                  paste0("D2d8"),
                  paste0("D2d9"),
                  paste0("D2d10"),
                  paste0("D2d11"),
                  paste0("D2d12"),
                  paste0("D2d13"),
                  paste0("D2d14"),
                  paste0("D2d15"),
                  paste0("D2d16"),
                  paste0("D2d17"),
                  paste0("D2d18"),
                  paste0("D2d21"),
                  paste0("D2d26"),
                  paste0("D2d28"),
                  paste0("D2d30"),
                  paste0("D2d32"),
                  paste0("D2d35"),
                  paste0("D2d42"),
                  paste0("D2d49"),
                  paste0("D2d57"),
                  paste0("D2d65"),
                  paste0("D2d74"),
                  paste0("D2d81"),
                  paste0("D2d87"))
save(file = "/home/wburr/Rushali_REST/Donor2/all_D2_PCA_noIS.rda", all_d)
rm(list = ls())
load("/home/wburr/Rushali_REST/Donor2/all_D2_PCA_noIS.rda")
```


## Cleanup Rushali Noted Compounds Throughout

There are a number of compounds that have to do with the
column, the reference standard (somehow coming through), 
and other definitely not-related things. We will strip
them out now, in preparation for PCA.

### Remove Explicitly

The following compounds are definitely not related to decomp,
and are related to the process or environment, and need to be
removed:

* Bromobenzene (or Benzene, bromo) 
* Oxygen
* Acetone
* Methyl alcohol / Methanol
* Carbon dioxide

Before we start that, we'll make a list of the actual
"appearances" of compounds, and then check against this
to determine the actual filtering arguments (e.g., first 8 characters
or full name, etc.).

```{r}
all_compounds <- sort(unique(
  unlist(lapply(all_d, FUN = function(x) { unlist(x$Name) }))))
length(all_compounds)
```

So there are 979 unique compounds present across the
total samples, after Controlling (but with the 2.0 ratio
argument in place). Let's look for each of the above compounds:

```{r benz}
loc1 <- grep(pattern = "Benzene, bromo-", all_compounds, 
             ignore.case = TRUE)
# all_compounds[loc1]
```

```{r oxy}
loc2 <- grep(pattern = "oxygen", all_compounds, ignore.case = TRUE)
# all_compounds[loc2]
```

```{r acetone}
if("Acetone" %in% all_compounds) {
  loc3 <- which("Acetone" %in% all_compounds)
} else {
  loc3 <- NULL
}
# all_compounds[loc3]
```

```{r methyl}
loc4 <- which("Methyl Alcohol" %in% all_compounds)
# all_compounds[loc4]
loc5 <- grep(pattern = "methanol, TMS", all_compounds, 
             ignore.case = TRUE)
# all_compounds[loc5]
```

```{r co2}
loc6 <- grep(pattern = "carbon dioxide", all_compounds, ignore.case = TRUE)
# all_compounds[loc6]
```

Put them together as indexes, then extract these from the
list of compounds as actual names.

```{r}
remove_specifics <- c(loc1, loc2, loc3, loc4, loc5, loc6)
remove1 <- all_compounds[remove_specifics]
```

### Remove Via Keyword

There are three keywords that show up that we should also
strip out:

* Sil
* TMS
* TBDMS

Let's grab these now:

```{r}
loc1 <- grep(pattern = "Sil", all_compounds, ignore.case = TRUE)
# all_compounds[loc1]
loc2 <- grep(pattern = "TMS", all_compounds)
# all_compounds[loc2]
loc3 <- grep(pattern = "TBDMS", all_compounds)
# all_compounds[loc3]
remove2 <- all_compounds[c(loc1, loc2, loc3)]
```

### Merge Compounds

We need some logic to look for things that are the same
compound, but different in only stereochemistry. The
indicator seems to be brackets: (E), (Z), (S),
popping up in one or more of the variants. So there might be,
for example, 2-Octene, (E)- as a compound, and then another
sample might have 2-Octene, (Z)-. These should just be merged. 

Let's try to look for them first:

```{r}
loc1 <- grep(pattern = "\\(E\\)-$", all_compounds)
loc2 <- grep(pattern = "\\(Z\\)-$", all_compounds)
loc3 <- grep(pattern = "\\(S\\)-$", all_compounds)
loc4 <- grep(pattern = "\\(R\\)-$", all_compounds)
to_clean <- c(loc1, loc2, loc3, loc4)
```

Now, the tricky bit: how to fix this up. What we want
are these compounds, and their corresponding compounds which
**don't** have the (S), (Z), (R) or (E); or have a different one.
In all cases, we'll merge them down to the **doesn't have brackets**
version if it exists, or if it doesn't, we'll make one.

```{r}
mappings <- data.frame(Original = NA, Transformed = NA)
for(j in 1:length(to_clean)) {
  orig <- all_compounds[to_clean[j]]
  fixed <- strsplit(orig, "\\(")[[1]][1]
  fixed <- substr(fixed, 1, nchar(fixed) - 2)
  mappings[j, ] <- c(orig, 
                     fixed)
}
```

## Back to the Original Data, Ready to Rock & Roll

So we have remove1 - compounds to remove. We have remove2 - 
more compounds to remove. And we have mappings, which have
compounds that need to be renamed. Then, at the end, we need
to check for duplicates, because the renaming may have 
resulted in more than one compound surviving in a single sample
due to the stereochemistry issue. 

```{r}
test <- lapply(all_d, FUN = function(x) { 
    y <- x %>% filter(!(Name %in% remove1 | Name %in% remove2))
    which_rows <- which(y$Name %in% mappings$Original)
    if(length(which_rows) > 0) {
      for(j in 1:length(which_rows)) {
        orig <- unlist(y[which_rows[j], "Name"])
        y[which_rows[j], "Name"] <- mappings[mappings$Original == orig,
                                             "Transformed"]
      }
    }
    y
  })
```

Now, look for duplicates, and remove if any now exist:

```{r}
test <- lapply(test, FUN = function(x) { 
    dupes <- which(duplicated(x$Name))
    if(length(dupes) > 0) {
      x[-dupes, ]
    } else {
      x
    }
  })
```

We're done! All fixed up. Let's write this back out to an Excel
file for Rushali to take a look at.

```{r}
library("xlsx")
names(test)[1]
write.xlsx(x = test[[1]],
           file = "/home/wburr/Rushali_REST/Donor2/all_D2_cleaned_noIS.xlsx",
           sheetName = names(test)[1], 
           col.names = TRUE,
           row.names = TRUE,
           append = FALSE)
for(j in 2:length(test)) {
  write.xlsx(x = test[[j]],
             file = "/home/wburr/Rushali_REST/Donor2/all_D2_cleaned_noIS.xlsx",
             sheetName = names(test)[j],
             col.names = TRUE,
             row.names = TRUE,
             append = TRUE) 
}
```


## Back to PCA

Now, the objective is to create "blanked" vectors, with 0s inserted in the
spots where detection was not found. This requires a full list of 
unique compounds, then we PCA away.

```{r}
unique_compounds <- sort(unique(
      unlist(lapply(test, FUN = function(x) { x$Name }))))
```

Then, let's create a set of vectors, 31 in total.

```{r}
pca_dat <- as.data.frame(matrix(data = 0.0, nrow = length(unique_compounds),
                                ncol = 32))
pca_dat[, 1] <- unique_compounds
names(pca_dat) <- c("Name", names(test))
for(j in 1:length(test)) {
  x <- test[[j]]
  x_names <- x$Name
  x_area <- apply(x, MAR = 1, FUN = function(y) { 
    z <- y[seq(2, length(y), 2)]; 
    z <- as.numeric(z);
    mean(z) })
  # manually loop through, so we don't get rearrangements ...
  for(k in 1:length(x_names)) {
    pca_dat[pca_dat$Name == x_names[k], j+1] <- x_area[k]
  }
  #pca_dat[pca_dat$Name %in% x_names, j+1] <- x_area
}
extracting <- pca_dat[, -1]
extracting <- apply(extracting, MAR = 2, FUN = function(x) { as.numeric(x) })
row.names(extracting) <- pca_dat$Name
pca_dat <- extracting

write.csv(file = "pca_ready_D2_noIS.csv", pca_dat)
```

